---
title: "ReadMe and Figures"
author: "Hengxing Zou"
date: "5/27/2022"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
source("Functions_SpatialBH.R")

fig_dir = "Figures"
if (!dir.exists(fig_dir)) {
  dir.create(fig_dir)
}

```

This document contains code needed to produce all figures in the main text and supporting information of "Mechanisms of Priority Effects Determine Dispersal-Diversity Relationships in Seasonal Metacommunities", as well as a brief readme section. All necessary functions are annotated in the `R` file `Functions_SpatialBH.R`. Before running following code, please make sure that the `R` file and this file are under the same directory.

# Basic Description of the Model

Local population dynamics are described by a modified Beverton-Holt model. Before dispersal and end-of-season mortality, the population at the end of the season $N_i'(t+1)$ is:

$$

N_{i}'(t+1) = \frac{\lambda_iN_i(t)}{1+\Sigma^n_{j=1}\alpha_{ij}\lambda_{i}N_{j}(t)}

$$

where population of the next season is determined by density-dependent population growth and the proportion of population surviving into the next generation. $N_i(t)$ and $N_j(t)$ denote population of this season but after dispersal and end-of-season mortality. Trait-mediated priority effects is allowed by calculating interspecific competition coefficients, $\alpha_{ij}$, as a function of relative time of emergence between two species, $\Delta p_{ij}$:

$$

\alpha_{ij} = \frac{B}{1+\exp({(d-\Delta p_{ij})/c})}

$$

where $\Delta p_{ij}=p_{j}-p_{i}$.

# Relaxing Assumptions in the Main Text

Simulation results of the main text are based on assumptions that each season contains only one generation, dispersal happens once at the end of each season, all species have equal dispersal rates, and no patch undergo extinction. These assumptions can be relaxed by adjusting relevant parameters of the simulation model. Although we do not provide simulation results of these specific situations, we provide full extension in the model that allows for these relaxations.

## Season Length

In seasonal communities, an end-of-season mortality is imposed by the environment to all species. In the main text, we use a simplified model that assumes only one generation per season ($T=1$). Under this assumption, there is no overlapping generations, and the survival $s$ is equivalent to end-of-season survival. However, we design our model to allow for multiple generations per season (as in multivoltine insects; $T>1$). In this case the inter-generational survival ($s$) and end-of-season survival are not the same, the latter requiring a new parameter, $s_e$. Like $s$, we assume that all species have the same $s_e$. In simulations, we keep these values separate, even in scenarios with one generation per season.

## Frequency of Dispersal

We assume dispersal happens only once at the end of the season. In nature, dispersal could happen multiple times throughout the season. We designate this frequency with an integer $f$, and only consider the case of more frequent dispersal when season length is larger than 1 (i.e. multiple generations per season). We fix one dispersal event at the end of the season, and randomly choose $f-1$ time points within the season for the rest of dispersal events. This requires $f-1<T$. Mid-season dispersal follows the same process described in the main text, characterized by dispersal rates $r$ and patches with dispersal $D$. Our model also enable one dispersal event per multiple seasons. In this case, the number of seasons without dispersal ($f_s$) is larger than 1, meaning that dispersal will only happen after $f_s$ seasons, and there will be no dispersal within these seasons.

## Patch Extinction

We assume that no patch within the metacommunity undergo extinction. Our model can simulate the situation when we relax this assumption by setting an additional parameter, number of patches with disturbance ($P_d$), to a nonzero integer. $P_d$ should not be larger than $P$, the number of patches in the metacommunity. At the end of the season, we randomly choose $P_d$ patches and set population of all species in these patches to 0. 

# Data Generation

**WARNING**

The following section takes a while to run: generating all data would take up to 5 hours, depending on the machine. We suggest using the provided `.RData` file for fast reproduction of figures in the main text. Note that because of the stochastic process these reproduced figures may not match exactly with figures in the main text.

## Initial conditions

- Parameters

The following chunk defines necessary parameters for running the simulation. Note that their names could be different from name listed in the main text. 

```{r}

nspp = 5 # number of species; n
size = 50 # number of patches; P
init_pop = 10 # initial population; N_0

intra = rep(0.07, nspp) # intraspecific competition coefficients of all speciesp; alpha_ii
B = 0.225 # asymptotic maximum of the competition-phenology function (Eqn. 2)
scal = 0.5 # midpoint of the competition-phenology function (Eqn. 2)

max_time = 0.5*(nspp-1) # maximum difference in time of emergence; equally partitioned by number of species to 
                        # generate average time of emergence for each species
var = 2 # seasonal variation of emergence times

lambda = rep(100, nspp) # intrinsic growth rates
season_length = 1 # number of generations in a season; T
n_seasons = 110 # length of simulation

disp_freq = 1 # dispersal frequency withint a season; f
disp_freq_seasons = 1 # dispersal frequency over seasons; f_s

Pd = 0 # number of patches with disturbance; P_d

```

- Initial Metacommunities

The following chunk initiate metacommunities according to the two scenarios outlined in the main text: 

Scenario 1: Initial populations are equal across all patches (Equal Initials);
Scenario 2: One species colonizes each patch first (First Colonizer).

We suggest reusing these initial metacommunities for all analyses because they are randomly generated and could lead to different results for each run. 

```{r}

M_1 = init_metacom(size, nspp, 10)
M_2 = init_metacom(size, nspp, 10, each = T)

```

We have also supplied initial metacommunities for running our simulations in the main text. To use these, make sure the folder `SavedMetacoms` are under the same directory as this file.

```{r}

M_1 = read.csv("SavedMetacoms/metacom_1.csv") %>% as.matrix()
M_2 = read.csv("SavedMetacoms/metacom_2.csv") %>% as.matrix()

```

## Population Dynamics

The following chunk provide code for simulating a time series of 100 seasons under each case and parameter combinations. The total number of cases is 8, with a factorial design of scenarios (Equal Initials or First Colonizer) $x$ presence of trait-mediated priority effects $x$ dispersal rates (low or high). For analyses using other intraspecific competition coefficients (Figure S8-S13), change the `intra` values below accordingly. After simulations, make sure to reset `intra` to the default value (0.07) before running other simulations.

**ESTIMATED RUNTIME: 30 MINS**

```{r}

# Low intra (shown in the main text)

intra = rep(0.07, nspp)

# Medium intra (shown in Appendix II, Figure S8-1S0)

# intra = rep(0.10, nspp)

# High intra (shown in Appendix II, Figure S11-S13)

# intra = rep(0.15, nspp)

# Low dispersal, Scenario 1

disp_rates = rep(0.01, nspp)

num_lowdisp_1 = multiple_reps(M_1, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                            lambda, intra, B, scal, 0, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_num_lowdisp_1 = met_df_multirep(num_lowdisp_1[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "num")

tra_lowdisp_1 = multiple_reps(M_1, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                            lambda, intra, B, scal, 1, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_tra_lowdisp_1 = met_df_multirep(tra_lowdisp_1[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "tra")

lowdisp_1 = rbind(dynamics_num_lowdisp_1, dynamics_tra_lowdisp_1) %>% mutate(disp_rate = 0.01)

# High dispersal, Scenario 1

disp_rates = rep(0.1, nspp)

num_highdisp_1 = multiple_reps(M_1, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                             lambda, intra, B, scal, 0, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_num_highdisp_1 = met_df_multirep(num_highdisp_1[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "num")

tra_highdisp_1 = multiple_reps(M_1, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                             lambda, intra, B, scal, 1, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_tra_highdisp_1 = met_df_multirep(tra_highdisp_1[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "tra")

highdisp_1 = rbind(dynamics_num_highdisp_1, dynamics_tra_highdisp_1) %>% mutate(disp_rate = 0.1)

# Low dispersal, Scenario 2

disp_rates = rep(0.01, nspp)

num_lowdisp_2 = multiple_reps(M_2, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                              lambda, intra, B, scal, 0, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_num_lowdisp_2 = met_df_multirep(num_lowdisp_2[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "num")

tra_lowdisp_2 = multiple_reps(M_2, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                              lambda, intra, B, scal, 1, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_tra_lowdisp_2 = met_df_multirep(tra_lowdisp_2[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "tra")

lowdisp_2 = rbind(dynamics_num_lowdisp_2, dynamics_tra_lowdisp_2) %>% mutate(disp_rate = 0.01)

# High dispersal, Scenario 2

disp_rates = rep(0.1, nspp)

num_highdisp_2 = multiple_reps(M_2, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                               lambda, intra, B, scal, 0, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_num_highdisp_2 = met_df_multirep(num_highdisp_2[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "num")

tra_highdisp_2 = multiple_reps(M_2, season_length, n_seasons, disp_rates, 50, disp_freq, disp_freq_seasons,
                               lambda, intra, B, scal, 1, max_time, var, rep = 50, n_disturb = Pd, burnin = 10)
dynamics_tra_highdisp_2 = met_df_multirep(tra_highdisp_2[[1]]) %>% filter(!(time == 0)) %>% mutate(time = time) %>% mutate(type = "tra")

highdisp_2 = rbind(dynamics_num_highdisp_2, dynamics_tra_highdisp_2) %>% mutate(disp_rate = 0.1)

```

## Dispersal-Diversity Relationships

The following chunks produce figures of dispersal-diversity relationships under two initial scenarios. For analyses using other intraspecific competition coefficients (Figure S10, S13), change the `intra` values below accordingly. After simulations, make sure to reset `intra` to the default value (0.07) before running other simulations.

**ESTIMATED RUNTIME: 60 MINS**

```{r}

# Low intra (shown in the main text)

intra = rep(0.07, 5)

# Medium intra (shown in Appendix II, Figure S8-S10)

# intra = rep(0.10, 5)

# High intra (shown in Appendix II, Figure S11-S13)

# intra = rep(0.15, 5)

disp_rates_lst = c(0.001, 0.005, seq(0.01, 0.1, 0.01))

# Scenario 1

num_Sce1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50, burnin = 10)
tra_Sce1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50, burnin = 10)

num_Sce1 %<>% mutate(type = "num")
tra_Sce1 %<>% mutate(type = "tra")
all_Sce1 = rbind(num_Sce1, tra_Sce1)

# Scenario 2

num_Sce2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50)
tra_Sce2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50)

num_Sce2 %<>% mutate(type = "num")
tra_Sce2 %<>% mutate(type = "tra")
all_Sce2 = rbind(num_Sce2, tra_Sce2)# %>% pivot_longer(cols = 2:4, names_to = "metric", values_to = "values")

```

## Different Variations in Phenology

**ESTIMATED RUNNING TIME: 180 MINS**

- Scenario 1

```{r}

# Low var

var = 0.5
num_lowvar_1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50, burnin = 10)
tra_lowvar_1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50, burnin = 10)

num_lowvar_1 %<>% mutate(type = "num")
tra_lowvar_1 %<>% mutate(type = "tra")
all_lowvar_1 = rbind(num_lowvar_1, tra_lowvar_1)

# Medium var

var = 2
num_midvar_1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50, burnin = 10)
tra_midvar_1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50, burnin = 10)

num_midvar_1 %<>% mutate(type = "num")
tra_midvar_1 %<>% mutate(type = "tra")
all_midvar_1 = rbind(num_midvar_1, tra_midvar_1)

# High var

var = 8
num_highvar_1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50, burnin = 10)
tra_highvar_1 = multiple_disp(M_1, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50, burnin = 10)

num_highvar_1 %<>% mutate(type = "num")
tra_highvar_1 %<>% mutate(type = "tra")
all_highvar_1 = rbind(num_highvar_1, tra_highvar_1)

```

- Scenario 2

```{r}

# Scenario 2

# Low var

var = 0.5
num_lowvar_2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50)
tra_lowvar_2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50)

num_lowvar_2 %<>% mutate(type = "num")
tra_lowvar_2 %<>% mutate(type = "tra")
all_lowvar_2 = rbind(num_lowvar_2, tra_lowvar_2)

# Medium var

var = 2
num_midvar_2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50)
tra_midvar_2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50)

num_midvar_2 %<>% mutate(type = "num")
tra_midvar_2 %<>% mutate(type = "tra")
all_midvar_2 = rbind(num_midvar_2, tra_midvar_2)

# High var

var = 8
num_highvar_2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 0, max_time, var = var, rep = 50)
tra_highvar_2 = multiple_disp(M_2, season_length, n_seasons, disp_rates_lst, c(50), disp_freq, disp_freq_seasons,
                         lambda, intra, B, scal, 1, max_time, var = var, rep = 50)

num_highvar_2 %<>% mutate(type = "num")
tra_highvar_2 %<>% mutate(type = "tra")
all_highvar_2 = rbind(num_highvar_2, tra_highvar_2)

```

# Figures in the Main Text

## Population Dynamics

- Plotting

```{r}

# Scenario 1

# p_pop_1 =
rbind(lowdisp_1, highdisp_1) %>% filter(patch == 5) %>% 
  ggplot(aes(x = time, y = log(population+1), color = species, linetype = rep)) + geom_line(alpha = 0.3) +
  theme_bw() + xlab("Season") + ylab("log(Population+1)") + 
  scale_color_manual(values = color_scheme) +
  scale_linetype_manual(values = rep(1, 50)) +
  facet_grid(disp_rate ~ type, scales = "fixed",             
             labeller = labeller(disp_rate = c(`0.01` = "Low r", `0.1` = "High r"),
                                 type = c(num = "No trait-mediated", tra = "Trait-mediated"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.position = "none",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

# Scenario 2

# p_pop_2 =
rbind(lowdisp_2, highdisp_2) %>% filter(patch == 5) %>% 
  ggplot(aes(x = time, y = log(population+1), color = species, linetype = rep)) + geom_line(alpha = 0.3) +
  theme_bw() + xlab("Season") + ylab("log(Population+1)") + 
  scale_color_manual(values = color_scheme) +
  scale_linetype_manual(values = rep(1, 50)) +
  facet_grid(disp_rate ~ type, scales = "fixed",             
             labeller = labeller(disp_rate = c(`0.01` = "Low r", `0.1` = "High r"),
                                 type = c(num = "No trait-mediated", tra = "Trait-mediated"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.position = "none",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

```

```{r, eval = F}

ggsave("Figures/Fig3.pdf", plot = p_pop_1, width = 2700, height = 1800, units = "px")
ggsave("Figures/Fig4.pdf", plot = p_pop_2, width = 2700, height = 1800, units = "px")

# If working with medium intraspecific competition
# ggsave("Figures/FigS8.png", plot = p_pop_1, width = 2700, height = 1800, units = "px")
# ggsave("Figures/FigS9.png", plot = p_pop_2, width = 2700, height = 1800, units = "px")

# If working with high intraspecific competition
# ggsave("Figures/FigS11.png", plot = p_pop_2, width = 2700, height = 1800, units = "px")
# ggsave("Figures/FigS12.png", plot = p_pop_2, width = 2700, height = 1800, units = "px")

```

## Dispersal-Diversity Relationships

- Plotting

```{r}

# p_dispdiv =
rbind(mutate(all_Sce1, init_cond = "equal_pop"), 
      mutate(all_Sce2, init_cond = "single_spp")) %>% 
  pivot_longer(cols = 2:4, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = disp_rate, y = value, color = type, fill = type)) + geom_smooth() +
  theme_bw() + xlab("Dispersal rate") + ylab("Value") + 
  scale_x_continuous(breaks = c(0, 0.05, 0.1), labels = c("0", "0.05", "0.1")) +
  scale_color_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) +
  scale_fill_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) + 
  facet_grid(metric ~ init_cond, scales = "free_y", 
             labeller = labeller(metric = c(alpha = "Alpha (Simpson's)", beta = "Beta (Bray-Curtis)", gamma = "Gamma (Simpson's)"),
                                 init_cond = c(equal_pop = "1. Equal Initials", single_spp = "2. First Colonizer"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_blank(), legend.position = "bottom",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

```

```{r, eval = F}

ggsave("Figures/Fig5.pdf", plot = p_dispdiv, width = 1800, height = 2400, units = "px")

# If working with medium intraspecific competition
# ggsave("Figures/FigS9.png", plot = p_dispdiv, width = 1800, height = 2400, units = "px")

# If working with high intraspecific competition
# ggsave("Figures/FigS12.png", plot = p_dispdiv, width = 1800, height = 2400, units = "px")

# If working with 10 species
# ggsave("Figures/FigS5.png", plot = p_dispdiv, width = 1800, height = 2400, units = "px")

```

- Plotting for Presentation

```{r}

all_Sce2 %>% 
  filter(type == "tra") %>%
  ggplot(aes(x = disp_rate, y = gamma)) + 
  geom_smooth(method = "loess", color = "#5A72D1", fill = "#5A72D1") +
  geom_jitter(color = "#5A72D1") +
  theme_bw() + 
  ylim(0, 1) +
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title = element_blank(), 
        panel.grid.minor = element_blank())

```

# Figures in the Appendix

## Final Occupancy

The following chunk plots final occupancy of the dominant species in each repetition, defined as the proportion of patches in which the regionally dominant species has the highest population; the regional dominant species is determined as the species that has the highest regional population for each repetition. This takes a while.

```{r}

all_pop_dynamics = rbind(lowdisp_1 %>% filter(time == 110) %>% mutate(init_cond = "equal_pop"), 
                         highdisp_1 %>% filter(time == 110) %>% mutate(init_cond = "equal_pop"), 
                         lowdisp_2 %>% filter(time == 110) %>% mutate(init_cond = "single_spp"), 
                         highdisp_2 %>% filter(time == 110) %>% mutate(init_cond = "single_spp"))

# Determine the species with highest total population (dominant species)

dom_spp = all_pop_dynamics %>% group_by(species, rep, type, disp_rate, init_cond) %>% 
  summarize(population = sum(population)) %>% ungroup() %>% 
  group_by(rep, type, disp_rate, init_cond) %>% filter(population == max(population)) %>% 
  mutate(species = as.integer(species))

# Species that has the highest population in each patch

max_pop = all_pop_dynamics %>% 
  group_by(patch, rep, type, disp_rate, init_cond) %>% 
  filter(population == max(population))

# Count how many times the dominant species is the species with the highest population in each repetition

all_counts = data.frame()

for (r in 1:50) {
  subset = dom_spp %>% filter(rep == r) # all dominant species
  for (i in 1:nrow(subset)) {
    spp_count = max_pop %>% filter(rep == r, type == subset[i, ]$type, 
                       disp_rate == subset[i, ]$disp_rate, init_cond == subset[i, ]$init_cond)
    spp = subset[i, ]$species # dominant species of this particular parameter combination
    counted = spp_count %>% count(species)
    occupancy = filter(counted, species == spp)$n/sum(counted$n)
    all_counts = rbind(all_counts, data.frame(rep = r, type = subset[i, ]$type, 
                                              disp_rate = subset[i, ]$disp_rate, init_cond = subset[i, ]$init_cond,
      occupancy = occupancy))
  }
}

# Calculate mean and standard deviation of occupancy data

all_counts_summ = all_counts %>% group_by(type, disp_rate, init_cond) %>% 
  summarize(mean = mean(occupancy), sd = sd(occupancy))

# Plotting

p_dom =
ggplot() + 
  geom_point(data = all_counts, aes(x = as.factor(disp_rate), y = occupancy, color = type), 
              alpha = 0.3, position = position_jitterdodge(jitter.width = 0.2)) +
  geom_errorbar(data = all_counts_summ,
                aes(x = as.factor(disp_rate), ymin = mean-sd, ymax = mean+sd, color = type),
                width = 0, position = position_dodge(0.75)) +
  geom_point(data = all_counts_summ, 
             aes(x = as.factor(disp_rate), y = mean, fill = type), 
             shape = 21, color = "black", size = 3, position = position_dodge(0.75)) + 
  theme_bw() + xlab("Dispersal rate") + ylab("Occupancy of the\nregionally most \nabundant species") + #ylim(0, 1) +
  scale_x_discrete(label = c("Low", "High")) +
  ylim(0, 1) +
  scale_color_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) + 
  scale_fill_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) + 
  facet_grid(. ~ init_cond, scales = "free_y", 
             labeller = labeller(init_cond = c(equal_pop = "1. Equal Initials", single_spp = "2. First Colonizer"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_blank(), legend.position = "bottom",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

```

```{r, eval = F}

ggsave("Figures/FigS1.png", plot = p_dom, width = 1800, height = 1800, units = "px")

```

## Temporal Turnover

The following chunks calculate and plots temporal turnover of metacommunity over fixed intervals using methods of Legendre (2019). The plotted dissimilarity index is the $D$ in Legendre 2019, which describes the collective effects of species gains and losses between two time points, excluding the first 10 seasons as burn-in. These chunks produce Figure 5.

These chunks use data generated by the [Population Dynamics] section.

```{r}

tb_low_num1 = temporal_beta(num_lowdisp_1[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "num", disp_rate = 0.01)
tb_low_tra1 = temporal_beta(tra_lowdisp_1[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "tra", disp_rate = 0.01)
tb_high_num1 = temporal_beta(num_highdisp_1[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "num", disp_rate = 0.1)
tb_high_tra1 = temporal_beta(tra_highdisp_1[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "tra", disp_rate = 0.1)

tb_Sce1 = rbind(tb_low_num1, tb_low_tra1, tb_high_num1, tb_high_tra1) %>% mutate(init_cond = "single_spp")

tb_low_num2 = temporal_beta(num_lowdisp_2[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "num", disp_rate = 0.01)
tb_low_tra2 = temporal_beta(tra_lowdisp_2[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "tra", disp_rate = 0.01)
tb_high_num2 = temporal_beta(num_highdisp_2[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "num", disp_rate = 0.1)
tb_high_tra2 = temporal_beta(tra_highdisp_2[[1]], step = 10) %>% filter(time > 20) %>% mutate(type = "tra", disp_rate = 0.1)

tb_Sce2 = rbind(tb_low_num2, tb_low_tra2, tb_high_num2, tb_high_tra2) %>% mutate(init_cond = "equal_pop")

# Plotting

# p_tempdis = 
rbind(tb_Sce1, tb_Sce2) %>% #mutate(init_cond = fct_relevel(init_cond, "single_spp", "equal_pop")) %>% 
  ggplot(aes(x = time, y = D, color = type, fill = type)) + geom_smooth() + 
  theme_bw() + xlab("Season") + ylab("Dissimilarity index") + ylim(0, 0.8) + 
  scale_color_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) +
  scale_fill_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) +
  facet_grid(disp_rate ~ init_cond, scales = "free_y", 
             labeller = labeller(disp_rate = c(`0.01` = "Low r", `0.1` = "High r"),
                                 init_cond = c(equal_pop = "1. Equal Initials", single_spp = "2. First Colonizer"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_blank(), legend.position = "bottom",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

```

```{r, eval = F}

ggsave("Figures/FigS2.png", plot = p_tempdis, width = 1800, height = 1800, units = "px")

```

## Population at Different Times

The following chunks calculate population of each species of one randomly selected patch at different times of the simulation. 

```{r}

# Scenario 1

# p_poptime_1 = 
rbind(lowdisp_1, highdisp_1) %>% filter(patch == 5 & 
                                    (time == 35 | time == 60 | time == 85 | time == 110)) %>% #season_length*n_seasons+10
  ggplot(aes(x = as.factor(time), y = log(population+1), fill = species)) + geom_boxplot() + 
  theme_bw() + xlab("Season") + ylab("log(Population+1)") +
  scale_fill_manual(values = color_scheme, guide = "none") +
  facet_grid(disp_rate ~ type, scales = "fixed",             
             labeller = labeller(disp_rate = c(`0.01` = "Low r", `0.1` = "High r"),
                                 type = c(num = "No trait-mediated", tra = "Trait-mediated"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.position = "none",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

# Scenario 2

# p_poptime_2 = 
rbind(lowdisp_2, highdisp_2) %>% filter(patch == 5 & 
                                    (time == 35 | time == 60 | time == 85 | time == 110)) %>% #season_length*n_seasons+10
  ggplot(aes(x = as.factor(time), y = log(population+1), fill = species)) + geom_boxplot() + theme_bw() +
  xlab("Season") + ylab("log(Population+1)") +
  scale_fill_manual(values = color_scheme, guide = "none") + 
  facet_grid(disp_rate ~ type, scales = "fixed",             
             labeller = labeller(disp_rate = c(`0.01` = "Low r", `0.1` = "High r"),
                                 type = c(num = "No trait-mediated", tra = "Trait-mediated"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.position = "none",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

```

```{r, eval = F}

ggsave("Figures/FigS3.png", plot = p_poptime_1, width = 2700, height = 1800, units = "px")
ggsave("Figures/FigS4.png", plot = p_poptime_2, width = 2700, height = 1800, units = "px")

```

## Effects of Variation

The following chunks generate figures that show how dispersal-diversity relationships are affected by the magnitude of seasonal variations.

```{r}

# Scenario 1

# p_ddvar_1 = 
rbind(mutate(all_lowvar_1, var = as.factor(0.5)), 
      mutate(all_midvar_1, var = as.factor(2)), 
      mutate(all_highvar_1, var = as.factor(8))) %>%
  pivot_longer(cols = 2:4, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = disp_rate, y = value, color = type, fill = type)) + geom_smooth() +
  theme_bw() + xlab("Dispersal rate") + ylab("Value") + 
  scale_x_continuous(breaks = c(0, 0.05, 0.1), labels = c("0", "0.05", "0.1")) +
  scale_color_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) +
  scale_fill_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) + 
  facet_grid(metric ~ var, scales = "free_y", 
             labeller = labeller(metric = c(alpha = "Alpha (Simpson's)", beta = "Beta (Bray-Curtis)", gamma = "Gamma (Simpson's)"),
                                 var = c(`0.5` = "Small", `2` = "Medium", `8` = "Large"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_blank(), legend.position = "bottom",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

# Scenario 2

# p_ddvar_2 = 
rbind(mutate(all_lowvar_2, var = as.factor(0.5)), 
      mutate(all_midvar_2, var = as.factor(2)), 
      mutate(all_highvar_2, var = as.factor(8))) %>%
  pivot_longer(cols = 2:4, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = disp_rate, y = value, color = type, fill = type)) + geom_smooth() +
  theme_bw() + xlab("Dispersal rate") + ylab("Value") + 
  scale_x_continuous(breaks = c(0, 0.05, 0.1), labels = c("0", "0.05", "0.1")) +
  scale_color_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) +
  scale_fill_manual(values = c("#5A72D1", "#D97437"), labels = c("No trait-mediated", "Trait-mediated")) + 
  facet_grid(metric ~ var, scales = "free_y", 
             labeller = labeller(metric = c(alpha = "Alpha (Simpson's)", beta = "Beta (Bray-Curtis)", gamma = "Gamma (Simpson's)"),
                                 var = c(`0.5` = "Small", `2` = "Medium", `8` = "Large"))) + 
  theme(axis.text.y = element_text(size = 15), axis.text.x = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_blank(), legend.position = "bottom",
        panel.grid.minor = element_blank(), 
        panel.spacing.x = unit(1, "lines"), panel.spacing.y = unit(0.5, "lines"),
        strip.text.x = element_text(size = 15), strip.text.y = element_text(size = 15))

```

```{r, eval = F}

ggsave("Figures/FigS6.png", plot = p_ddvar_1, width = 2400, height = 2400, units = "px")
ggsave("Figures/FigS7.png", plot = p_ddvar_2, width = 2400, height = 2400, units = "px")

```
